A new tidy data structure to support exploration and modeling of temporal data

### 2019-11-10 reviews/response.Rmd

---
title: Response to reviewers
author: JCGS-19-044.R1
output:
  MonashEBSTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

We thank the reviewers for their comments which has improved this paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.

@@

# Reviewer 1 comments {-}

> Bottom of page 8: "Figure 3 peels the tidy module with" I'm not sure about the use of the word peels there?

We rephrased "peels" to "demonstrates".

> Top of page 9: "The rigidity of tsibble, as the fundamental data infrastructure, warrants the validity of temporal data analysis for the later stage." Should "warrants" be replaced with "ensures" and perhaps replace "temporal data analysis for the later stage" with "subsequent temporal data analysis".

@@

Done.

> P6 L34. Second item missing full stop at the end.

Done.

> I think for Figure 4, you should mention that supplementary figure contains the gif if the animation does not work or perhaps include a link to a working animation.

Thanks. We added a sentence in Figure 4 caption "The animations are available with the supplementary materials online, and can be viewed at <https://github.com/earowang/paper-tsibble/blob/master/img/animate-1.gif>."

> P15 L58. Mention `future_*` is from `furrr` package.

Done. We also mentioned the **future** package.

> P16 L24. Missing reference for `lubridate`.

### 2019-11-10 tsibble.Rmd
minor revisions

In R [@R-base], time series and longitudinal data are of different representations. The native `ts` object and the enhancements of **zoo** [@zoo2005] and **xts** [@R-xts], assemble time series into wide matrices with implicit time indexes. If there are multiple sub-groups, such as country or product type, these would be kept in different data objects. A relatively new R package **tibbletime** [@R-tibbletime] proposed a data class of *time tibble* to represent time series in heterogeneous long format. It only requires an index variable to be declared. However, this is insufficient, and a more rigid data structure is required for temporal analytics and modeling. The **plm** [@plm2008] and **panelr** [@R-panelr] packages both manage longitudinal data in long format.

@@

1. Index is a variable with inherent ordering from past to present.
2. Key is a set of variables that define observational units over time
3. Each observation should be uniquely identified by index and key.
4. Each observational unit should be measured at a common interval, if regularly spaced.

@@

Inspired by a "primary key" [@codd_relational_1970], a unique identifier for each observation in a relational database, the tsibble key also uniquely identifies each observational unit over time. When constructing a tsibble, any duplicates of key-index (i.e. key-value) pairs will fail, because duplicates signal a data quality issue, which would likely affect subsequent analyses and hence decision making. For example, either `gender` or `country` alone is not enough to be the key for the tuberculosis data. Analysts are encouraged to better understand the data, or reason about the process of data cleaning when handling duplicates. Figure \@ref(fig:tsibble-tidy) peels the tidy module with clear routes required for a tsibble. The rigidity of tsibble, as the fundamental data infrastructure, warrants the validity of temporal data analysis for the later stage.

@@

(ref:animate-cap) An illustration of a window of size 5 to compute rolling averages over annual tuberculosis cases in Australia using sliding, tiling and stretching. (Animation needs to be viewed with Adobe Acrobat Reader.)

@@

Rolling windows not only iterate but roll over a sequence of elements of a fixed window. A complete and consistent set of tools are available for facilitating window-related operations, a family of `slide()`, `tile()`, `stretch()`, and their variants. `slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs. For type stability, the functions always return lists. Other variants including `*_lgl()`, `*_int()`, `*_dbl()`, `*_chr()` return vectors of the corresponding types, as well as `*_dfr()` and `*_dfc()` for row-binding and column-binding data frames respectively. Their multiprocessing equivalents prefixed by `future_*()` enable rolling in parallel [@R-future; @R-furrr].

@@

All user-facing functions can be roughly organized into three major chunks according to their functionality: vector functions (1d), table verbs (2d), and window family. Each chunk is an independent module, but works interdependently. Vector functions in the package mostly operate on time. The atomic functions (such as `yearmonth()` and `yearquarter()`) can be embedded in the `index_by()` verb to collapse a tsibble to a less granular interval. Since they are not tied to a tsibble, they can be used in a broader range of data applications not constrained to tsibble. On the other hand, the table verbs can incorporate many other vector functions from a third party, like the **lubridate** package.

>>>>

In R [@R-base], time series and longitudinal data are of different representations. The native `ts` object and the enhancements by **zoo** [@zoo2005] and **xts** [@R-xts], assemble time series into wide matrices with implicit time indexes. If there are multiple sub-groups, such as country or product type, these would be kept in different data objects. A relatively new R package **tibbletime** [@R-tibbletime] proposed a data class of *time tibble* to represent time series in heterogeneous long format. It only requires an index variable to be declared. However, this is insufficient, and a more rigid data structure is required for temporal analytics and modeling. The **plm** [@plm2008] and **panelr** [@R-panelr] packages both manage longitudinal data in long format.

@@

1. Index is a variable with inherent ordering from past to present.
2. Key is a set of variables that define observational units over time.
3. Each observation should be uniquely identified by index and key.
4. Each observational unit should be measured at a common interval, if regularly spaced.

@@

Inspired by a "primary key" [@codd_relational_1970], a unique identifier for each observation in a relational database, the tsibble key also uniquely identifies each observational unit over time. When constructing a tsibble, any duplicates of key-index (i.e. key-value) pairs will fail, because duplicates signal a data quality issue, which would likely affect subsequent analyses and hence decision making. For example, either `gender` or `country` alone is not enough to be the key for the tuberculosis data. Analysts are encouraged to better understand the data, or reason about the process of data cleaning when handling duplicates. Figure \@ref(fig:tsibble-tidy) demonstrates the tidy module with clear routes required for a tsibble. The rigidity of tsibble, as the fundamental data infrastructure, ensures the validity of subsequent temporal data analysis.

@@

(ref:animate-cap) An illustration of a window of size 5 to compute rolling averages over annual tuberculosis cases in Australia using sliding, tiling and stretching. The animations are available with the supplementary materials online, and can be viewed at <https://github.com/earowang/paper-tsibble/blob/master/img/animate-1.gif>.

@@

Rolling windows not only iterate but roll over a sequence of elements of a fixed window. A complete and consistent set of tools are available for facilitating window-related operations, a family of `slide()`, `tile()`, `stretch()`, and their variants. `slide()` expects one input, `slide2()` two inputs, and `pslide()` multiple inputs. For type stability, the functions always return lists. Other variants including `*_lgl()`, `*_int()`, `*_dbl()`, `*_chr()` return vectors of the corresponding types, as well as `*_dfr()` and `*_dfc()` for row-binding and column-binding data frames respectively. Their multiprocessing equivalents prefixed by `future_*()` enable rolling in parallel, via **future** [@R-future] and **furrr** [@R-furrr].

@@

All user-facing functions can be roughly organized into three major chunks according to their functionality: vector functions (1d), table verbs (2d), and window family. Each chunk is an independent module, but works interdependently. Vector functions in the package mostly operate on time. The atomic functions (such as `yearmonth()` and `yearquarter()`) can be embedded in the `index_by()` verb to collapse a tsibble to a less granular interval. Since they are not tied to a tsibble, they can be used in a broader range of data applications not constrained to tsibble. On the other hand, the table verbs can incorporate many other vector functions from a third party, like the **lubridate** package [@R-lubridate].
